### 2018-04-30 papers/synth/conclusion.tex
intro
encoding finished

\ashu{Discuss the model in Biological terms.}

>>>>

In this paper, we presented the encoding of the synthesis problems
that may rise from VTSs.
%
We demonstrated that our tool scale up to the relevant sizes of the
VTSs. 
%
We will take this tool to the biologists and develop wet experiments 
that may validate some synthesis results from the tool.
%
Our model of VTSs are static graphs.
%
In future, we will study the dynamic behaviours of VTSs.
%
It will allow us to predict behaviors after the perturbations in the VTSs
and more ways to test the predicted synthesis results.

### 2018-04-30 papers/synth/encoding.tex
encoding finished

% % We will consider several variations of the problem.
% %
% \begin{itemize}
% \item Learning activity functions
% \item Modifying labels of the nodes and edges
% \item Adding missing edges
% \item Adding 
% \end{itemize}

%Types of questions we are trying to solve.

%
%
\begin{enumerate}
\item  Add new edges 
\item  Modify the flow of molecules on the nodes and on the edges.
  \begin{enumerate}
  \item use available set of molecules.
  \item allow new molecule type to be added.     
  \end{enumerate}        
\item Learning meaningful activity functions from a given language of
  propositional formulas
  % \begin{enumerate}
  %   \item kcnf: low depth circuit.
  %   \item Boolean gates: And, Or.
  %   \item Boolean gates with linear combination.  
  % \end{enumerate}        
  % - Function dependence with var occurring once.
\item  VTS repair. Allow both delete and add to the system.
\end{enumerate}

@@

In our synthesis algorithm, we take a VTS $G =
(\nodes,\mols,\edges,\nlabel,\pairs,\edgef,\nodef)$ as input.
%
We allow activity functions not to be specified, i.e., vectors
$\edgef$ and $\nodef$ may contain uninterpreted Boolean functions.
%
We construct the following constraints to encode the available information
about $G$.
%
\begin{align*}
  \texttt{KnownE} &= \land \{e_{i,j,q,m}|(i,M_1,j),...,(i,M_{q'},j) \in \edges \land q \leq q' \land m \in M_q \}\\
  \texttt{KnownN} &= \land \{n_{i,m}| m \in \nlabel(i) \land i \in \nodes \}\\
  \texttt{KnownP} &= \land \{p_{i,j}| (i,j) \in \pairs \} \\
  \texttt{KnownActiveN} &= \land \{ a_{i,m} = \nodef_m (n_{i,1},\dots,n_{i,\mu}) | \nodef_m \text{ is defined.} \} \\
  \texttt{KnownActiveE} &= \land \{ b_{i,j,q,m} = \edgef_m(e_{i,j,q,1}, .., e_{i,j,q,\mu})
                   | \edgef_m \text{ is defined.} \} \\
  \texttt{KnownCons} & = \texttt{KnownE} \land \texttt{KnownN} \land 
 \texttt{KnownP} \land \texttt{KnownActiveE} \land \texttt{KnownActiveN}
\end{align*}
We also collect the variables that are absent in the VTS.
\begin{align*}
  \texttt{UnknownELabel} &=
  \{e_{i,j,q,m}|(i,M_1,j),...,(i,M_{q'},j) \in \edges \land 
                          0 < q \leq q' \land m \not\in M_{q} \}\\
  \texttt{UnknownE} &= \{e_{i,j,q}|(i,M_1,j),...,(i,M_{q'},j) \in \edges \land 
                    q' < q \leq \pi \}\\
  \texttt{UnknownNLabel} &= \{n_{i,m}| m \not\in \nlabel(i) \land i \in \nodes \}\\
  \texttt{UnknownP} &= \{p_{i,j}| (i,j) \not\in \pairs \} \\
  \texttt{UnnownActive} &=  \Land \{ a_{i,m} = \nodef_m (n_{i,1},\dots,n_{i,\mu}) | \nodef_m \text{ is undefined.} \} \union \\
   &\quad\quad\quad  \{ b_{i,j,q,m} = \edgef_m(e_{i,j,q,1}, .., e_{i,j,q,\mu})
                   | \edgef_m \text{ is undefined.} \}
\end{align*}
We have defined $\texttt{UnknownE}$, $\texttt{UnknownN}$, and
$\texttt{UnknownP}$
as set and will be converted into formulas depending
on the different usage in the synthesis problems.

@@

Now we will consider the encodings for  the above
listed synthesis problems.
%
The variations are presentations are to represent all the encodings supported by our tool.
%
However, the combinations of the variation are also possible and our tool easily
supports them.
%
For simplicity of the presentation, we assume that if we are synthesizing an aspect of VTS,
then all other aspects are fully given.

\subsubsection{Fixing VTS by adding edges}
%
Now we will consider the case when we add new edges to VTS to satisfy the properties.
%
In the following, the pseudo-Boolean formula $\texttt{AddE}$ encodes
that at most $slimit$ new undeclared edges may be added in VTS.
%
\begin{align*}
  \texttt{AddE}(slimit) &= \Sigma~\texttt{UnknownE} \leq slimit\\
  \texttt{FixedE} &= \texttt{UnknownActive} \land \lnot \Lor \texttt{UnknownELabel} \;\union\;
                    \texttt{UnknownNLabel} \;\union\;
                    \texttt{UnknownP}
\end{align*}
We put together the constraints so far and obtain the following formula.
\begin{align*}
  & \texttt{SynthE}(slimit) = \\
  & \quad\quad (  \texttt{Consistancy}\land \texttt{Property} \land
  \texttt{KnownCons} \land \texttt{FixedE} \land
  \texttt{AddE}(slimit) )
  % \tag{SynthE}\label{eq:addedge1}
\end{align*}
As we have seen $\texttt{Consistancy}$ encodes the basic constraints about VTS,
$\texttt{Property}$ encodes the goal, and $\texttt{KnownCons}$ are the constraints
that are fixed in VTS.
%
A satisfying model of $\texttt{SynthE}$ will represent the synthesized VTS.
%
In the later synthesis problems, we will construct a similar QBF formula with same
first three parts and the last one is due the requirements of the synthesis problem.

@@

\paragraph{\bf Fixing VTS by adding molecules to the labels:}
The system may also be fixed only by modifying labels on the edges or the nodes instead
of adding edges.
%
In the following, the formula encodes that only $slimit$ edge labels may be added.
\begin{align*}
  &\texttt{AddLabelEdge}(slimit) = 
    \Sigma~\texttt{UnknwonEdgeLabel}  \leq slimit\\
  & \texttt{FixedForLabel} = \texttt{UnknownActive} \land \lnot \Lor \texttt{UnknownE} \;\union\;
                    \texttt{UnknownNLabel} \;\union\;
                    \texttt{UnknownP}\\
  &\texttt{SynthLabel}(slimit) = \\
  & \quad\quad
    (  \texttt{Consistancy}\land \texttt{Property} \land
  \texttt{KnownCons} \land \texttt{FixedForLabel} \land \texttt{AddLabelEdge}(slimit) )
\end{align*}
Similar to the previous encoding we solve the satisfiability of the above formula
to obtain additional molecules that may be added to the edge labels.

@@

The formula \texttt{NNFTemplate} encodes a set of negation normal form functions 
that take $y_1,..,y_\mu$ as input and contain $\lambda$ literals.
To encode the set, it also has finite-range integer variables
$z_1,..,z_{2\lambda}$ as parameters to select the functions.

\begin{align*}
  \texttt{Gate}( z, w_1, w_2 ) = ( z  = 1 \limplies w_1  \land w_2 ) \land  
  ( z = 2  \limplies w_1  \lor w_2 ) \land
  ( z = 3  \limplies w_1 )\\
  \texttt{Leaf}( z, [y_1,..,y_\mu] ) =
  \Land\limits_{l=1}^{\mu} ( z = 2l-1  \limplies y_{l}) \land ( z = 2l  \limplies \lnot y_{l})\\
  \texttt{NNFTemplate}([z_1,..,z_{2\lambda}],[y_1,..,y_\mu] ) = \\
  \exists w_1,..,w_{2\lambda}.\;w_1 \land \Land_{l=1}^{\lambda} w_{\lambda+l} = \texttt{Leaf}( z_{\lambda+l}, [y_1,..,y_\mu] ) \land
  w_{l} = \texttt{Gate}( z_l, w_{2l},w_{2l+1})
\end{align*}

Given the above templates

\begin{align*}
  &\texttt{FindFunctions}(z,\lambda) =\\
  &\Land \{ \bigwedge\limits_{i} n_{i,m} \limplies a_{i,m} = 
  \texttt{NNFTemplate}([z_{m,1},..,z_{m,2\lambda}],[n_{i,1},\dots,n_{i,\mu}] ) |
  \nodef_m \text{ is undefined}\} \\
  &\Land \{ \bigwedge\limits_{i,j,q} e_{i,j,q,m} \limplies b_{i,j,q,m} = 
  \texttt{NNFTemplate}([z_{i,j,q,1},..,z_{i,j,q,2\lambda}],[e_{i,j,q,1},..,e_{i,j,q,\mu}] ) | 
  \edgef_m \text{ is undefined}
  \}
\end{align*}

\begin{align*}
    & \texttt{FixedForLabel} = \lnot \Lor \texttt{UnknownE} \;\union\; \texttt{UnknownN} \;\union\;
                    \texttt{UnknownNLabel} \;\union\;
                    \texttt{UnknownP}\\
  &\texttt{SynthFunction}(z,\lambda) = \\
  & \quad\quad
    (  \texttt{Consistancy}\land \texttt{Property} \land
  \texttt{KnownCons} \land \texttt{FixedForFunctions} \land \texttt{FindFunctions}(slimit) )
\end{align*}

\subsection{Repairing VTS:}
%
Now we will consider the repair of VTS by allowing not only addition but also
deletion of the molecules, edges, functions, or pairing matrix.
%
We illustrate the repairing on one class of variables and rest can be .....
%
The repair of the VTS requires both the addition and the deletion of
the present molecules, node molecules pairing matrix and activity of
the molecules. We have encoded the vts repair in our tool using the
dummy xor bits.

>>>>

\begin{enumerate}
\item  Add new edges 
\item  Modify the flow of molecules on the nodes and on the edges.
  % \begin{enumerate}
  % \item use available set of molecules.
  % \item allow new molecule type to be added.     
  % \end{enumerate}        
\item Learning meaningful activity functions from a given language of
  propositional formulas
  % \begin{enumerate}
  %   \item kcnf: low depth circuit.
  %   \item Boolean gates: And, Or.
  %   \item Boolean gates with linear combination.  
  % \end{enumerate}        
  % - Function dependence with var occurring once.
\item  VTS repair. Allow both delete and add to the system.
\end{enumerate}

@@

In our synthesis method, we take a VTS $G =
(\nodes,\mols,\edges,\nlabel,\pairs,\edgef,\nodef)$ as input.
%
We allow activity functions not to be specified, i.e., vectors
$\edgef$ and $\nodef$ may contain uninterpreted Boolean functions.
%
We construct the following constraints to encode the available information
about $G$.
%
\begin{align*}
  \texttt{KnownE} &= \land \{e_{i,j,q,m}|(i,M_1,j),...,(i,M_{q'},j) \in \edges \land q \leq q' \land m \in M_q \}\\
  \texttt{KnownN} &= \land \{n_{i,m}| m \in \nlabel(i) \land i \in \nodes \}\\
  \texttt{KnownP} &= \land \{p_{i,j}| (i,j) \in \pairs \} \\
  \texttt{KnownActiveN} &= \land \{ a_{i,m} = \nodef_m (n_{i,1},\dots,n_{i,\mu}) | \nodef_m \text{ is defined.} \} \\
  \texttt{KnownActiveE} &= \land \{ b_{i,j,q,m} = \edgef_m(e_{i,j,q,1}, .., e_{i,j,q,\mu})
                   | \edgef_m \text{ is defined.} \} \\
  \texttt{KnownCons} & = \texttt{KnownE} \land \texttt{KnownN} \land 
 \texttt{KnownP} \land \texttt{KnownActiveE} \land \texttt{KnownActiveN}
\end{align*}
We also collect the variables that are not set to true in $\texttt{KnownCons}$.
\begin{align*}
  \texttt{UnknownELabel} &=
  \{e_{i,j,q,m}|(i,M_1,j),...,(i,M_{q'},j) \in \edges \land 
                          0 < q \leq q' \land m \not\in M_{q} \}\\
  \texttt{UnknownE} &= \{e_{i,j,q}|(i,M_1,j),...,(i,M_{q'},j) \in \edges \land 
                    q' < q \leq \pi \}\\
  \texttt{UnknownNLabel} &= \{n_{i,m}| m \not\in \nlabel(i) \land i \in \nodes \}\\
  \texttt{UnknownP} &= \{p_{i,j}| (i,j) \not\in \pairs \} \\
  \texttt{UnnownActive} &=  \Land \{ a_{i,m} = \nodef_m (n_{i,1},\dots,n_{i,\mu}) | \nodef_m \text{ is undefined.} \} \union \\
   &\quad\quad\quad  \{ b_{i,j,q,m} = \edgef_m(e_{i,j,q,1}, .., e_{i,j,q,\mu})
                   | \edgef_m \text{ is undefined.} \}
\end{align*}
We have defined $\texttt{UnknownE}$, $\texttt{UnknownN}$, and
$\texttt{UnknownP}$
as set and will be converted into formulas depending
on the different usage in the synthesis problems.

@@

Now we will consider the encodings for  the above
listed synthesis problems.
%
The presented variations represent the encodings
supported by our tool.
%
Additionally, the combinations of the variation are also possible and
our tool easily supports them.
%
For simplicity of the presentation, we assume that if we are
synthesizing an aspect of VTS, then all other aspects are fully given.
%
Therefore, we will describe two kinds of constraints for synthesis
problems.
%
The first will encode the variable part in the synthesis problem and
the other encodes the fixed parts.
%
Subsequently, the two constraints will be put together with 
$\texttt{Consistancy}$ and 
$\texttt{Property}$.

\subsubsection{Fixing VTS by adding edges}
%
Now we will consider the case when we add new edges to VTS to satisfy the properties.
%
In the following, the pseudo-Boolean formula $\texttt{AddE}$ encodes
that at most $slimit$ new undeclared edges may be added in the VTS.
%
\texttt{FixedForEdge} encodes the parts of the VTS that do not change.
\begin{align*}
  &\texttt{AddE}(slimit) = \sum~\texttt{UnknownE} \leq slimit\\
  &\texttt{FixedForEdge} = \texttt{KnownCons} \land \texttt{UnknownActive} \land\\
   & \quad\quad\quad\quad \lnot \Lor \texttt{UnknownELabel} \;\union\;
                    \texttt{UnknownNLabel} \;\union\;
                    \texttt{UnknownP}
\end{align*}
We put together the constraints and obtain the following formula.
\begin{align*}
  & \texttt{SynthE}(slimit) =
     (  \texttt{Consistancy}\land \texttt{Property} \land
   \texttt{FixedForEdge} \land
  \texttt{AddE}(slimit) )
  % \tag{SynthE}\label{eq:addedge1}
\end{align*}
As we have seen $\texttt{Consistancy}$ encodes the basic constraints about VTS,
$\texttt{Property}$ encodes the goal, and
the rest two are just defined constraints.
%
A satisfying model of $\texttt{SynthE}$ will represent the synthesized VTS.
%
In the later synthesis problems, we will construct a similar QBF
formula with same first two parts and the last two are due the
requirements of the synthesis problem.

@@

\paragraph{\bf Fixing VTS by adding molecules to the labels:}
The system may also be fixed only by modifying labels on the edges or the nodes instead
of adding edges.
%
In the following, the formula encodes that only $slimit$ edge labels may be added.
\begin{align*}
  &\texttt{AddLabelEdge}(slimit) = 
    \sum~\texttt{UnknwonEdgeLabel}  \leq slimit\\
  & \texttt{FixedForLabel} = \texttt{KnownCons} \land \texttt{UnknownActive} \land\\
  & \quad \quad \lnot \Lor \texttt{UnknownE} \;\union\;
                    \texttt{UnknownNLabel} \;\union\;
                    \texttt{UnknownP}\\
  &\texttt{SynthLabel}(slimit) = \\
  & \quad\quad
    (  \texttt{Consistancy}\land \texttt{Property} \land
   \texttt{FixedForLabel} \land \texttt{AddLabelEdge}(slimit) )
\end{align*} 
Similar to the previous encoding we solve the
satisfiability of the above formula to obtain additional molecules
that may be added to the node labels.

In the following, the formula \texttt{NNFTemplate} encodes a set of
negation normal form functions that take $y_1,..,y_\mu$ as input and
contain $\lambda$ literals.
%
We use $\texttt{Gate}$ to encode a gate that takes a parameter integer
$x$ to encode various gates.
%
We use $\texttt{Leaf} $ to encode the literal at some position.
%
Both are stitched to define \texttt{NNFTemplate}.
%
To encode the set of NNF formulas with $\lambda$ literals, it has
finite-range integer variables
$z_1,..,z_{2\lambda}$ as parameters.

\begin{align*}
  &\texttt{Gate}( x, w_1, w_2 ) = ( x  = 1 \limplies w_1  \land w_2 ) \land  
  ( x = 2  \limplies w_1  \lor w_2 ) \land
  ( x = 3  \limplies w_1 ) \land ...\\
  &\texttt{Leaf}( x, [y_1,..,y_\mu] ) =
  \Land\limits_{l=1}^{\mu} ( x = 2l-1  \limplies y_{l}) \land ( x = 2l  \limplies \lnot y_{l})\\
  &\texttt{NNFTemplate}([z_1,..,z_{2\lambda}],[y_1,..,y_\mu] ) = \\
  &\exists w_1,..,w_{2\lambda}.\;w_1 \land \Land_{l=1}^{\lambda} w_{\lambda+l} = \texttt{Leaf}( z_{\lambda+l}, [y_1,..,y_\mu] ) \land
  w_{l} = \texttt{Gate}( z_l, w_{2l},w_{2l+1})
\end{align*}

Using the template we define the constraints $\texttt{FindFunctions}(z,\lambda)$
that encodes the candidate functions that satisfy the activity requirements,
where $z$ is the vectors of parameters for encoding parameters
for each molecule,
and $\lambda$ limits the size of the candidate functions. 
%
We fix the all other aspect of the VTS to be fixed via constraints
$\texttt{FixedForFunctions}$.
%
\begin{align*}
  &\texttt{FindFunctions}(z,\lambda) =\\
  &\Land \{ \bigwedge\limits_{i} n_{i,m} \limplies a_{i,m} = 
  \texttt{NNFTemplate}([z_{m,1},..,z_{m,2\lambda}],[n_{i,1},\dots,n_{i,\mu}] ) |
  \nodef_m \text{ is undefined}\} \\
  &\Land \{ \bigwedge\limits_{i,j,q} e_{i,j,q,m} \limplies b_{i,j,q,m} = 
  \texttt{NNFTemplate}([z_{i,j,q,1},..,z_{i,j,q,2\lambda}],[e_{i,j,q,1},..,e_{i,j,q,\mu}] ) | 
  \edgef_m \text{ is undefined}
  \}\\
  & \texttt{FixedForFunctions} = \texttt{KnownCons} \land\\
  & \quad \quad
\lnot \Lor \texttt{UnknownE} \;\union\; \texttt{UnknownELabel} \;\union\;
                    \texttt{UnknownNLabel} \;\union\;
                    \texttt{UnknownP}\\
  &\texttt{SynthFunction}(z,\lambda) = \\
  & \quad\quad
    (  \texttt{Consistancy}\land \texttt{Property} \land
   \texttt{FixedForFunctions} \land \texttt{FindFunctions}(slimit) )
\end{align*}
We construct $\texttt{SynthFunction}(z,\lambda)$ similar to the earlier
variations.
%
By reading of the values of $z$ in a satisfying model of the formula,
we learn the synthesized function.

\subsection{Repairing VTS:}
%
Now we will consider the repair of VTS by allowing not only addition but also
deletion of the molecules, edges, functions, or pairing matrix.
%
We have encoded the vts repair in our tool by introducing flip bits
for each variable that is modifiable in the VTS.
%
We illustrate the repairing on one class of variables and rest can be
trivially extended.
%
Let us consider repairing of node labels.
%
For each bit $n_{i,m}$, we create a bit $flip_{i,m}$.
%
We add constraints that take xor of VTS assigned values for  $n_{i,m}$
and $flip_{i,m}$.
%
We also limit how many times the $flip_{i,m}$ can be true, therefore
limiting the number of flips.
%
Similar to the earlier variations, we construct
$\texttt{SynthRepairNode}(slimit)$ for the repair.
% 
\begin{align*}
    &\texttt{FlipN}(slimit) = \Land \{n_{i,m}\lxor flip_{i,m}| m \in \nlabel(i) \land i \in \nodes \} \land \\
  & \quad \quad \quad
    \Land \{ \lnot n_{i,m}\lxor flip_{i,m}| m \not \in \nlabel(i) \land i \in \nodes \} \land  \sum\limits_{i,m} flip_{i,m} \leq slimit \\
% \\  
    & \texttt{FixedForNodeRepair} =  \texttt{KnownE} \land  
      \texttt{KnownP} \land \texttt{KnownActiveE} \land \texttt{KnownActiveN}
      \land \\
  & \quad \quad \texttt{UnknownActive} \land \lnot 
    \Lor \texttt{UnknownE} \;\union\; \texttt{UnknownNLabel} \;\union\;
    \texttt{UnknownP}\\
  &\texttt{SynthRepairNode}(slimit) = \\
  & \quad\quad
    (  \texttt{Consistancy}\land \texttt{Property} \land
    \texttt{FixedForNodeRepair} \land \texttt{FlipN}(slimit) )
\end{align*}

@@
